Supplementary figures

### 2015-07-21 Manuscript/manuscript.tex
Making vocabulary more consistent and clear

To address the above questions, we developed a stochastic patch occupancy model (SPOM) in which each patch can be in one of three possible states: occupied by susceptible hosts (S), occupied by infectious hosts (I), and unoccupied by the host ($\emptyset$).  State transitions are governed by host colonization, extinction, and infection rates, where a susceptible population can become infectious either through direct contact with infectious immigrants, or through a local environmental pathogen reservoir (Table~\ref{transitions}).

We assume that the host population size supported by a patch is proportional to its quality ($A_i$), and that disease-induced mortality reduces the size of infectious populations by a factor of $\nu$.  Building on the framework developed by \cite{Hanski2003} then, the rates at which patch \emph{i}, with quality $A_i$, is colonized by susceptible and infectious individuals, respectively, is given by its connectivity:
\begin{align}
C_{Si}(t) &= A_i^{\xi_{im}} \sum_{i\neq j }\phi_j(S, t)A_j^{\xi_{em}}\exp(-D d_{ij}),\\
C_{Ii}(t) &= A_i^{\xi_{im}} \sum_{i\neq j }\phi_j(I, t)(\nu A_j) ^{\xi_{em}}\exp(-D d_{ij}),
\label{connectivity}
\end{align}

where $\phi_j(X, t)$ is an indicator function that is $1$ if patch $j$ is in state $X$ at time $t$ and is $0$ otherwise; $\xi_{im}$ and $\xi_{em}$ control how rates of immigration and emigration, respectively, scale with patch quality; $D$ is the inverse of the host's mean dispersal distance; and $d_{ij}$ is the effective distance between patches $i$ and $j$ .  Essentially, $C_{Xi}$ sums the colonization effort (or propagule pressure) from all patches in state $X$ to focal patch $i$.

@@

Susceptible populations are infected via direct contact with infectious colonists at rate $\delta C_{Ii}$, where $\delta$ is the direct transmission probability.  In addition, environmental transmission can take place when a susceptible population occupies a previously infected patch, and occurs at rate:
\begin{equation}
\gamma(\tau_i)=\gamma_0exp(-r\tau_{i}),
\end{equation}
where $\gamma_0$ is the initial infection rate of the pathogen reservoir, $\tau_{i}$ is the time since last infectious occupancy ($\tau_{i} = \infty$, and thus $\gamma(\tau_i) = 0$, if the patch has never been occupied by infectious hosts), and $r$ is the pathogen's decay rate in the environment (determined by the pathogen's environmental longevity).

@@

The model was parameterized according to the following assumptions (Table~\ref{params}). 
(1) The rate of emigration from a patch increases with population size, while the extinction rate of a patch decreases with populations size ($\xi_{im}=0.5$, $\alpha=1$, \cite{Hanski2003}).  
(2) Migrants preferentially select high quality habitat, that is, the rate at which a patch is colonized scales with quality ($\xi_{em}=0.5$, \cite{Hanski2003}, though we relax this below).  
(3)  Connectivity among patches is determined entirely by quality ($d_{ij}=1$ for all $i \neq j$).  
In addition to these assumptions, we chose parameters so that without infection, approximately 0.75 of the metapopulation was occupied ($\mu = 0.1$, $D=5$).  We varied disease induced mortality ($\nu$), probability of direct infection ($\delta$), and environmental longevity to explore the dynamics of a range of possible pathogens.

@@

To explore the relative influence of low and high quality habitat on host and pathogen dynamics, we simulated the spread of a range of pathogens in metapopulations with patch quality distributions shifted towards either high or low quality habitat.
These two distributions were chosen to have equal variance and an overlapping core of patches around unit quality, with either the minimum or maximum shifted to simulate regions with more, or less, high quality habitat (yielding distributions with qualities ranging from 0.23 to 1.24 and from 0.76 to 1.77).  
We then implemented a full-factorial set of simulations on each quality distribution in which infectious survival ($\nu$) and probability of direct infection ($\delta$) were each varied over 10 values (ranging from 0.1 to 1, and 0 to 0.9, respectively).  In addition, pathogen longevity was varied over 3 values (ranging from 0.3 to 3 times the expected residence time of a healthy population on a unit quality patch, i.e., 10 units of time).

For each simulation -- 100 replicates for each combination of disease parameters ($\nu$, $\delta$, and longevity), habitat quality distribution (high or low), and landscape structure (fully connected or lattice), qualities for 100 patches were drawn from a uniform distribution on the given habitat quality range (above). In each case, an entirely susceptible population was simulated until it reached an approximate steady state, at which point a randomly chosen occupied patch was infected.  The state of the metapopulation was then tracked for 5000 time steps.  We recorded the proportion of patches in each state at the end of each simulation and also computed the mean effective population size of susceptible and infectious hosts over the last 500 time steps of the simulation (assuming an effective population size of $A_i$ for patches that are occupied and susceptible and $\nu A_i$ for occupied infectious patches).

@@

A wide range of disease dynamics were observed over the range of parameters and habitat quality distributions explored.  
Epidemic outcome was influenced most strongly by infectious survival, with high infectious survival (i.e. low disease-induced mortality) facilitating widespread infection (in which no susceptible host populations persist), and very low infectious survival (i.e. high disease-induced mortality) facilitating disease-free dynamics or intermediate spread (Fig~\ref{endemic}). 
Pathogens with higher environmental longevities were generally able to invade and spread more easily, even with high disease-induced morality (Fig~\ref{endemic}(c, f)).  
The distribution of habitat quality in the metapopulation also played an important role in determining disease dynamics.  Increasing overall habitat quality facilitated widespread infection for a larger range of parameters (Fig~\ref{endemic}, a-c vs. d-f).
Metapopulations with a lattice structure exhibited qualitatively similar patterns, but with generally more limited pathogen spread and larger regions of susceptible maintenance (Supplement).

For a wide range of parameter values, high quality patch distributions had a net positive effect on total host population size relative to low quality distributions despite generally facilitating pathogen spread (red shaded regions in Fig~\ref{highvlow}).  
In the case of pathogens that were unable to invade due to low infectious survival and weak transmission, high quality habitat distributions supported larger host populations than low quality distributions due to the greater stability and connectivity afforded by high quality habitat (bottom left corner of the panels in Fig~\ref{highvlow}). 
Similarly, when high infectious survival and strong transmission (from either environmental or direct transmission) allowed a pathogen to spread widely (i.e. at a pandemic level) in both habitat quality distributions, high quality distributions again supported larger host populations than low quality distributions (top portion of panels in Fig~\ref{highvlow}).

However, for pathogens falling between these extremes (e.g. with intermediate disease-induced mortality), low quality habitat was a net benefit to the host, supporting larger total population sizes than high quality habitat (blue regions of Fig ~\ref{highvlow}).
The spread of these pathogens was strongly influenced by the habitat quality distribution, with low quality habitat limiting disease spread and better maintaining susceptible occupancy relative to high quality habitat (Fig~\ref{popsizes}, Supplemental).
In these cases, high quality habitat served as a sort of metapopulation-level trap, that is, otherwise beneficial habitat that is detrimental to the host population in the presence of disease.
As pathogen longevity increased, the range of the parameter space where we observed this metapopulation-level trap became smaller, apparent only for the most virulent pathogens.
Again, metapopulations with a lattice structure displayed similar patterns (Supplement).

@@

This expands on previous studies of disease spread in a homogeneous metapopulation that focused on movement rate as the driver of this trade-off, with increased movement both improving metapopulation stability and increasing disease spread (\cite{Hess1996}, \cite{Gog2002}, \cite{Park2012}).  
Shifting the patch quality distribution towards high quality habitat in our model is conceptually similar to increasing movement in these models.
And, as in these models, the net consequences of this trade-off for the host metapopulation depend on the characteristics of the disease.
Broadly, we can divide the observed dynamics into three different scenarios: limited, intermediate, and widespread infection (possibly leading to host extinction). 
High quality habitats are a net benefit to the host population when pathogen spread is either limited or is high enough that both high and low quality patches are affected by disease. 
In these cases, the connectivity and stability of high quality patches facilitate larger total population sizes with little additional pathogen spread.

There is an intermediate scenario, however, where low quality habitats are more likely to remain uninfected and as a result boost the metapopulation size while high quality patches that are heavily infected tend to reduce the metapopulation size. 
In this scenario, the ability of pathogens to invade and spread is influenced substantially by the colonization and extinction rates of high and low quality habitat, and the contrasting roles that these habitat types play in the metapopulation.

@@

Because high quality habitat supports larger host population sizes, these populations are less susceptible to environmental stochasticity and thus experience less frequent extinction events.
Due to this additional stability, high quality habitat is better able to both maintain a consistent pool of susceptible hosts to which infection can spread, and support infectious host populations under disease-induced mortality. 
As a result, the probability of infection (i.e. proportion of time occupied by infectious hosts) increases with patch quality (Supplement).
From the perspective of a migrant susceptible host then, selection of high quality habitat leads to increased disease risk relative to other habitats, either because the patch is likely already occupied by infectious hosts, or because the local susceptible hosts are likely to soon become infected (either through the presence of an environmental reservoir or direct contact).

@@

In contrast to high quality habitat, low quality patches help to limit pathogen spread and increase susceptible occupancy in the metapopulation (Fig~\ref{endemic}, \ref{popsizes}). 
Even though individual low quality patches, due to their high extinction and low colonization rates, are only able to support transient occupancy (Fig~\ref{simvis}, Supplement), the presence of low quality habitat facilitates more widespread susceptible persistence for a greater range of pathogens (Fig~\ref{endemic}).  
This phenomenon represents the other side of the role played by high quality patches in that low quality patches are relatively unstable (i.e. low population sizes and high extinction rates) and thus do not support infectious populations for long.
Since they are infrequently colonized (when high quality habitat is preferentially selected, $\xi_{em} > 0$), the environmental reservoir left behind by these infectious populations likely decays before it has the opportunity to infect new susceptible colonists. 
As a result, low quality patches effectively represent a dead-end for the pathogen, reducing the number of patches from which it can spread, and allowing susceptible hosts to more easily persist and avoid infection throughout the metapopulation.
The opportunity for pathogen reservoir decay provided by the ephemeral occupancy of low quality habitat echoes the 'migratory escape' hypothesis (\cite{Loehle1995}), which suggests that annual migration allows hosts to vacate contaminated sites and return once they are clean.
Taken with this literature, our results highlight the importance of transient/temporary habitat in mitigating host exposure to environmentally transmitted pathogens.

@@

\begin{figure}
\includegraphics[scale=0.4]{figure/figure_4.pdf}
\centering
\caption{Results from a single representative simulation with habitat quality uniformly distributed between 0.23 and 1.77, and a pathogen with a disease induced mortality of $\nu = 0.2$, a direct transmission rate of $\delta = 0.3$, and an envioronmental longevity equal to the expected residence time of a susceptible population on a unit quality patch (10 units of time here).  The top panel shows total proportion of patches occupied by susceptible (blue) and infectious (red) hosts over time.  The bottom panel shows the state -- susceptible (blue), infectious (red), or unoccupied (white) -- of individual patches through time.  Patches are stacked vertically with the lowest quality at the bottom and the highest quality at the top.  Note the overall more consistent occupancy on high quality habitat and the only very brief spurts of infectious occupancy on low quality habitat.}
\label{simvis}
\end{figure}

@@

\begin{table}
\caption{State transitions and their rates for patch $i$ in a metapopulation simulation.  $S$ denotes occupied by  susceptible hosts, $I$ denotes occupied infectious hosts, and $\emptyset$ denotes unoccupied by the host.  All patches are characterized by an environmental infection rate, $\gamma(\tau_i)$, where $\tau_i$ is the time since the patch was last occupied by infectious individuals.}
\begin{tabular}{llr}
State Transition & Process &  Rate \\
\hline
$S \rightarrow I$ & Infection (from contact or reservoir) & $\delta C_{Ii} + \gamma(\tau_i)$\\
$S \rightarrow \emptyset $ & Extinction (of susceptible) & $E_{Si}$\\
$I \rightarrow \emptyset $ & Extinction (of infectious) &  $E_{Ii}$ \\
$\emptyset \rightarrow S$ & Colonization (by susceptibles) & $C_{Si}$\\
$\emptyset \rightarrow I$ & Colonization (by infectious) & $C_{Ii}$\\
\label{transitions}
\end{tabular}
\end{table}

\begin{table}[h!]   
\caption{Parameters of the SPOM model, their meaning, and the values, or range of values, assigned.}
\begin{tabular}{p{1cm} p{6cm} r}
Param. & Interpretation &  Value(s) \\
\hline
$\xi_{im}$ & Effect of patch quality on immigration (preference for high quality habitat) & 0, 0.5 \\
$\xi_{em}$ & Effect of population size on emigration & 0.5  \\
$D$& Inverse of mean dispersal distance & 2 (lattice), 5 (full) \\
$d_{ij}$ & Distance between patch $i$ and $j$ & \parbox[t]{4cm}{\raggedleft $1 \forall i, j$ neighbors (lattice)\\ $1 \forall  i\neq j$ (full)}\\
\hline
$\mu$ & Extinction rate of unit quality patch & 0.1 \\
$\nu$ & Infectious survival & 0.1 - 1 \\
$\alpha$ & Strength of environmental stochasticity & 1 \\
\hline
$\delta$ & Probability of direct infection & 0 - 0.9 \\
$\gamma_0$ & Initial rate of infection from reservoir patch & 0.5
\end{tabular}
\label{params}
\end{table}

>>>>

To address the above questions, we developed a stochastic patch occupancy model (SPOM) in which each patch can be in one of three possible states: occupied by susceptible hosts (S), occupied by infectious hosts (I), and unoccupied by the host ($\emptyset$).  State transitions are governed by host colonization, extinction, and transmission rates, where a susceptible population can become infectious either through direct contact with infectious immigrants, or through a local environmental pathogen reservoir (Table~\ref{transitions}).

We assume that the host population size supported by a patch is proportional to its quality ($A_i$), and that a proportion, $\nu$, of hosts survive when a population is infected (i.e. the size of infectious populations is reduced by a factor of $\nu$).  Building on the framework developed by \cite{Hanski2003} then, the rates at which patch \emph{i}, with quality (or equivalently population size) $A_i$, is colonized by susceptible and infectious individuals, respectively, is given by its connectivity:
\begin{align}
C_{Si}(t) &= A_i^{\xi_{im}} \sum_{i\neq j }\phi_j(S, t)A_j^{\xi_{em}}\exp(-D d_{ij}),\\
C_{Ii}(t) &= A_i^{\xi_{im}} \sum_{i\neq j }\phi_j(I, t)(\nu A_j) ^{\xi_{em}}\exp(-D d_{ij}),
\label{connectivity}
\end{align}
where $\phi_j(X, t)$ is an indicator function that is $1$ if patch $j$ is in state $X$ at time $t$ and is $0$ otherwise; $\xi_{im}$ and $\xi_{em}$ control how rates of immigration and emigration, respectively, scale with patch quality/population size; $D$ is the inverse of the host's mean dispersal distance; and $d_{ij}$ is the effective distance between patches $i$ and $j$ .  Essentially, $C_{Xi}$ sums the colonization effort (or propagule pressure) from all patches in state $X$ to focal patch $i$.

@@

Susceptible populations are infected via direct contact with infectious colonists at rate $\delta C_{Ii}$, where $\delta$ is the direct transmission probability.  In addition, environmental transmission can take place when a susceptible population occupies a previously infected patch, and occurs at rate:
\begin{equation}
\gamma(\tau_i)=\gamma_0exp(-r\tau_{i}),
\end{equation}
where $\gamma_0$ is the initial transmission rate of the pathogen reservoir, $\tau_{i}$ is the time since last infectious occupancy ($\tau_{i} = \infty$, and thus $\gamma(\tau_i) = 0$, if the patch has never been occupied by infectious hosts), and $r$ is the pathogen's decay rate in the environment (determined by the pathogen's longevity, defined as its half-life in the environment).

@@

The model was parameterized according to the following assumptions (Table~\ref{params}). 
(1) The rate of emigration from a patch increases with the local population size, while the extinction rate of a patch decreases with populations size ($\xi_{im}=0.5$, $\alpha=1$, \cite{Hanski2003}).  
(2) Migrants preferentially select high quality habitat, that is, the rate at which a patch is colonized scales with quality ($\xi_{em}=0.5$, \cite{Hanski2003}, though we relax this below).  
(3)  Connectivity among patches is determined entirely by quality ($d_{ij}=1$ for all $i \neq j$).  
In addition to these assumptions, we chose parameters so that without infection, approximately 0.75 of the patches were occupied ($\mu = 0.1$, $D=5$).  We varied infectious survival ($\nu$), probability of direct transmission ($\delta$), and environmental longevity to explore the dynamics of a range of possible pathogens.

@@

To explore the relative influence of low and high quality habitat on host and pathogen dynamics, we simulated the spread of a range of pathogens in metapopulations with patch quality distributions shifted towards either high or low quality habitat.
These two distributions were chosen to have equal variance and an overlapping core of patches around unit quality, with either the minimum or maximum shifted to simulate regions with more, or less, high quality habitat (yielding distributions with qualities ranging from 0.23 to 1.24 and from 0.76 to 1.77).  
We then implemented a full-factorial set of simulations on each quality distribution in which infectious survival ($\nu$) and probability of direct transmission ($\delta$) were each varied over 10 values (ranging from 0.1 to 1, and 0 to 0.9, respectively).  In addition, pathogen longevity was varied over 3 values (ranging from 0.3 to 3 times the expected residence time of a healthy population on a unit quality patch, i.e., 10 units of time).

For each simulation -- 100 replicates for each combination of disease parameters ($\nu$, $\delta$, and longevity), habitat quality distribution (high or low), and landscape structure (fully connected or lattice), qualities for 100 patches were drawn from a uniform distribution on the given habitat quality range (above). In each case, an entirely susceptible population was simulated until it reached an approximate steady state, at which point a randomly chosen occupied patch was infected.  The state of the metapopulation was then tracked for 5000 time steps.  We recorded the proportion of patches in each state at the end of each simulation and also computed the mean total population size of susceptible and infectious hosts over the last 500 time steps of the simulation (summing the local population sizes for all occupied patches, assuming a local population size of $A_i$ for patches that are occupied and susceptible and $\nu A_i$ for occupied infectious patches).

@@

A wide range of disease dynamics were observed over the range of parameters and habitat quality distributions explored.  
Epidemic outcome was influenced most strongly by infectious survival, with high infectious survival (i.e. low disease-induced mortality) facilitating widespread infection (in which no susceptible host populations persist), and very low infectious survival (i.e. high disease-induced mortality) facilitating disease-free dynamics or intermediate spread (Fig~\ref{endemic}). 
Pathogens with higher environmental longevities were generally able to invade and spread more easily, even with low infectious survival (Fig~\ref{endemic}(c, f)).  
The distribution of habitat quality in the metapopulation also played an important role in determining disease dynamics.  Increasing overall habitat quality facilitated widespread infection for a larger range of parameters (Fig~\ref{endemic}, a-c vs. d-f).
Metapopulations with a lattice structure exhibited qualitatively similar patterns, but with generally more limited pathogen spread and larger regions of susceptible maintenance (Supplement).

For a wide range of parameter values, high quality patch distributions had a net positive effect on total host population size relative to low quality distributions despite generally facilitating pathogen spread (red shaded regions in Fig~\ref{highvlow}).  
In the case of pathogens that were unable to invade due to low infectious survival and weak transmission, high quality habitat distributions supported larger total host populations than low quality distributions due to the greater stability and connectivity afforded by high quality habitat (bottom left corner of the panels in Fig~\ref{highvlow}). 
Similarly, when high infectious survival and strong transmission (from either environmental or direct transmission) allowed a pathogen to spread widely (i.e. at a pandemic level) in both habitat quality distributions, high quality distributions again supported larger total host populations than low quality distributions (top portion of panels in Fig~\ref{highvlow}).

However, for pathogens falling between these extremes (e.g. with intermediate infectious survival), low quality habitat was a net benefit to the host, supporting larger total population sizes than high quality habitat (blue regions of Fig ~\ref{highvlow}).
The spread of these pathogens was strongly influenced by the habitat quality distribution, with low quality habitat limiting disease spread and better maintaining susceptible occupancy relative to high quality habitat (Fig~\ref{popsizes}, Supplemental).
In these cases, high quality habitat served as a sort of metapopulation-level trap, that is, otherwise beneficial habitat that is detrimental to the overall host population in the presence of disease.
As pathogen longevity increased, the range of the parameter space where we observed this metapopulation-level trap became smaller, apparent only for the most virulent pathogens.
Again, metapopulations with a lattice structure displayed similar patterns (Supplement).

@@

This expands on previous studies of disease spread in a homogeneous metapopulation that focused on movement rate as the driver of this trade-off, with increased movement both improving metapopulation stability and increasing disease spread (\cite{Hess1996}, \cite{Gog2002}, \cite{Park2012}).  
Shifting the patch quality distribution towards high quality habitat in our model is conceptually similar to increasing movement in these models.
And, as in these models, the net consequences of this trade-off for the host metapopulation depend on the characteristics of the disease.
Broadly, we can divide the observed dynamics into three different scenarios: limited, intermediate, and widespread infection (possibly leading to host extinction). 
High quality habitats are a net benefit to the total host population when pathogen spread is either limited or is high enough that both high and low quality patches are affected by disease. 
In these cases, the connectivity and stability of high quality patches facilitate larger total population sizes with little additional pathogen spread.

There is an intermediate scenario, however, where low quality habitats are more likely to remain uninfected and as a result boost the total population size while high quality patches that are heavily infected tend to reduce the total population size. 
In this scenario, the ability of pathogens to invade and spread is influenced substantially by the colonization and extinction rates of high and low quality habitat, and the contrasting roles that these habitat types play in the metapopulation.

@@

Because high quality habitat supports larger local host population sizes, these populations are less susceptible to environmental stochasticity and thus experience less frequent extinction events.
Due to this additional stability, high quality habitat is better able to both maintain a consistent pool of susceptible hosts to which infection can spread, and support infectious host populations under disease-induced mortality. 
As a result, the probability of infection (i.e. proportion of time occupied by infectious hosts) increases with patch quality (Supplement).
From the perspective of a migrant susceptible host then, selection of high quality habitat leads to increased disease risk relative to other habitats, either because the patch is likely already occupied by infectious hosts, or because the local susceptible hosts are likely to soon become infected (either through the presence of an environmental reservoir or direct contact).

@@

In contrast to high quality habitat, low quality patches help to limit pathogen spread and increase susceptible occupancy in the metapopulation (Fig~\ref{endemic}, \ref{popsizes}). 
Even though individual low quality patches, due to their high extinction and low colonization rates, are only able to support transient occupancy (Fig~\ref{simvis}, Supplement), the presence of low quality habitat facilitates more widespread susceptible persistence for a greater range of pathogens (Fig~\ref{endemic}).  
This phenomenon represents the other side of the role played by high quality patches in that low quality patches are relatively unstable (i.e. small local population sizes and high extinction rates) and thus do not support infectious populations for long.
Since they are infrequently colonized (when high quality habitat is preferentially selected, $\xi_{em} > 0$), the environmental reservoir left behind by these infectious populations likely decays before it has the opportunity to infect new susceptible colonists. 
As a result, low quality patches effectively represent a dead-end for the pathogen, reducing the number of patches from which it can spread, and allowing susceptible hosts to more easily persist and avoid infection throughout the metapopulation.
The opportunity for pathogen reservoir decay provided by the ephemeral occupancy of low quality habitat echoes the 'migratory escape' hypothesis (\cite{Loehle1995}), which suggests that annual migration allows hosts to vacate contaminated sites and return once they are clean.
Taken with this literature, our results highlight the importance of transient/temporary habitat in mitigating host exposure to environmentally transmitted pathogens.

@@

\begin{figure}
\includegraphics[scale=0.4]{figure/figure_4.pdf}
\centering
\caption{Results from a single representative simulation with habitat quality uniformly distributed between 0.23 and 1.77, and a pathogen with an infectious survival of $\nu = 0.2$, a direct transmission rate of $\delta = 0.3$, and an envioronmental longevity equal to the expected residence time of a susceptible population on a unit quality patch (10 units of time here).  The top panel shows total proportion of patches occupied by susceptible (blue) and infectious (red) hosts over time.  The bottom panel shows the state -- susceptible (blue), infectious (red), or unoccupied (white) -- of individual patches through time.  Patches are stacked vertically with the lowest quality at the bottom and the highest quality at the top.  Note the overall more consistent occupancy on high quality habitat and the only very brief spurts of infectious occupancy on low quality habitat.}
\label{simvis}
\end{figure}

@@

\begin{table}
\caption{State transitions and their rates for patch $i$ in a metapopulation simulation.  $S$ denotes occupied by  susceptible hosts, $I$ denotes occupied infectious hosts, and $\emptyset$ denotes unoccupied by the host.  All patches are characterized by an environmental transmission rate, $\gamma(\tau_i)$, where $\tau_i$ is the time since the patch was last occupied by infectious individuals.}
\begin{tabular}{llr}
State Transition & Process &  Rate \\
\hline
$S \rightarrow I$ & Transmission (from contact or reservoir) & $\delta C_{Ii} + \gamma(\tau_i)$\\
$S \rightarrow \emptyset $ & Extinction (of susceptible) & $E_{Si}$\\
$I \rightarrow \emptyset $ & Extinction (of infectious) &  $E_{Ii}$ \\
$\emptyset \rightarrow S$ & Colonization (by susceptibles) & $C_{Si}$\\
$\emptyset \rightarrow I$ & Colonization (by infectious) & $C_{Ii}$\\
\label{transitions}
\end{tabular}
\end{table}

\begin{table}[h!]   
\caption{Parameters of the SPOM model, their meaning, and the values, or range of values, assigned.}
\begin{tabular}{p{1cm} p{6cm} r}
Param. & Interpretation &  Value(s) \\
\hline
$\xi_{im}$ & Effect of patch quality on immigration (preference for high quality habitat) & 0, 0.5 \\
$\xi_{em}$ & Effect of population size on emigration & 0.5  \\
$D$& Inverse of mean dispersal distance & 2 (lattice), 5 (full) \\
$d_{ij}$ & Distance between patch $i$ and $j$ & \parbox[t]{4cm}{\raggedleft $1 \forall i, j$ neighbors (lattice)\\ $1 \forall  i\neq j$ (full)}\\
\hline
$\mu$ & Extinction rate of unit quality patch & 0.1 \\
$\nu$ & Infectious survival & 0.1 - 1 \\
$\alpha$ & Strength of environmental stochasticity & 1 \\
\hline
$\delta$ & Probability of direct transmission & 0 - 0.9 \\
$\gamma_0$ & Initial rate of transmission from reservoir patch & 0.5
\end{tabular}
\label{params}
\end{table}

### 2015-07-21 Manuscript/supplementary.tex
Making vocabulary more consistent and clear

\begin{figure}
\includegraphics[scale=0.5]{figure/supplement_2.pdf}
\centering
\caption{Differences between the median total effective population size of hosts on high quality patch distributions (quality between 0.76 and 1.77) and low quality patch distributions (quality between 0.23 to 1.24) as a function of disease-induced mortality ($\nu$) and the probability of direct transmission ($\delta$).  Reds indicate higher median host population sizes in high quality patch distributions, while blues indicate higher median host population sizes in low quality patch distributions.  Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.}
\label{highvlow}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{figure/supplement_3}
\centering
\caption{Probability of susceptible occupancy (measured as the proportion of time spent occupied by susceptible hosts over the last 500 time steps) as a function of patch quality. Red points indicate host preference for high quality habitat ($\xi_{im} = 0.5$), while black points indicate no preference ($\xi_{im} = 0$). Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.  Disease-induced mortality and probability of direct infection both fixed at $\nu = 0.2$ and $\delta = 0.3$, with fully connected landscape structure.}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{figure/supplement_4}
\centering
\caption{Probability of infectious occupancy (measured as the proportion of time spent occupied by infectious hosts over the last 500 time steps) as a function of patch quality. Red points indicate host preference for high quality habitat ($\xi_{im} = 0.5$), while black points indicate no preference ($\xi_{im} = 0$). Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.  Disease-induced mortality and probability of direct infection both fixed at $\nu = 0.2$ and $\delta = 0.3$, with fully connected landscape structure.}
\end{figure}

>>>>

\begin{figure}
\includegraphics[scale=0.5]{figure/supplement_2.pdf}
\centering
\caption{Differences between the median total effective population size of hosts on high quality patch distributions (quality between 0.76 and 1.77) and low quality patch distributions (quality between 0.23 to 1.24) as a function of infectious survival ($\nu$) and the probability of direct transmission ($\delta$).  Reds indicate higher median host population sizes in high quality patch distributions, while blues indicate higher median host population sizes in low quality patch distributions.  Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.}
\label{highvlow}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{figure/supplement_3}
\centering
\caption{Probability of susceptible occupancy (measured as the proportion of time spent occupied by susceptible hosts over the last 500 time steps) as a function of patch quality. Red points indicate host preference for high quality habitat ($\xi_{im} = 0.5$), while black points indicate no preference ($\xi_{im} = 0$). Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.  Infectious survival and probability of direct infection both fixed at $\nu = 0.2$ and $\delta = 0.3$, with fully connected landscape structure.}
\end{figure}

\begin{figure}
\includegraphics[scale=0.6]{figure/supplement_4}
\centering
\caption{Probability of infectious occupancy (measured as the proportion of time spent occupied by infectious hosts over the last 500 time steps) as a function of patch quality. Red points indicate host preference for high quality habitat ($\xi_{im} = 0.5$), while black points indicate no preference ($\xi_{im} = 0$). Panel columns show low (0.3, a), medium (1.0, b), and high (3.2, c) pathogen longevities.  Infectious survival and probability of direct infection both fixed at $\nu = 0.2$ and $\delta = 0.3$, with fully connected landscape structure.}
\end{figure}
